Classless Java: Tuning Java Interfaces

### 2015-12-08 paper/sections/Formalization.tex
S4.3 with comments

Defining \mBody{} is not trivial, and requires quite a lot of attention to the specific model of Java Interfaces, and on how it differs w.r.t. Java Class model.
$\mBody(\m,\C)$ denotes the actual method $\m$ (body included) that
interface $\C$ owns. The method can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.
\noindent$\begin{array}{l}
\InTextDef{15ex}{
\mBody(\m,\C_0)
}{
\override(\methods(\m),
\tops(\m,\Cs))
}\\
\InTextWith{\metaVar{IT}(\C_0) =
\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_n\ \oC\methods\ \cC \mbox{ and }
 \C\in\Cs \mbox{ if } \C_i\subtype\C, i\in 1..n

@@

\paragraph{\tops{}} recovers from the interface table only the ``needed'' methods, that is,
the non-static ones that are not reachable by another, less specific, superinterface.
Since the second parameter of \tops{} is a set, we can choose an arbitrary element to be $\C_0$.
In the definition we denote by $\Aux{originalMethod}(\m,\C)=\method$ the non-static method called $\m$ defined directly in $\C$.
%\hl{Another two functions $\super(\C_0)$ and $\get(\m,\C_0)$ are used for the definition. $\super(\C_0)$ collects the
%super interfaces of $\C_0$ recursively, and $\get(\m,\C_0)$ returns the non-static method declared in $\C_0$ with name $\m$, or $\emptyset$ otherwise.
%Finally, $\tops(\m,\C_0)$ returns the topmost methods $m$ from the super interfaces.}
Formally:

@@

%\paragraph{\shadow{}} chooses the most specific version of a method,
%that is the unique version available, or a conflicted version from a
%set of possibilities. In this case all the conflicting defintions are returned.
%We do not model overloading, so it is an error if multiple versions are available with different parameter types. Formally:
%
%\noindent
%$\begin{array}{l}
%\InTextDef{15ex}{
%\shadow()}{\none}\\
%\InTextDef{15ex}{\shadow(\method)}{\method}\\
%\InTextDef{15ex}{\shadow(\overline{\mh\QM;})}{\Aux{mostSpecific}(\overline{\mh\QM;})}\\
%\InTextDef{15ex}{\shadow(\methods)}{\methods}\\
%\InTextWith{\mbox{ the former case not applicable} }
%%\methods\mbox{ not of the form }\overline{\mh\QM;}
%%\mbox{ and }
%%\Aux{mostSpecific}(\methods)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}}\\
%\end{array}$
%
\paragraph{override} models how a method in an interface can
override an implementation in the superinterface, even in case of a
conflict.  Note how we use the special value $\none$, and how (forth
case) overriding can solve a conflict.
% The notation in the last case below
%means $\method'$ is not static, and is supertype of $\method$.

\noindent$\begin{array}{l}
\InTextDef{20ex}{\override(\none,\emptyset)}{\none}\\
\InTextDef{20ex}{\override(\method,\emptyset)}{\method}\\
\InTextDef{20ex}{\override(\none,\method)}{\method}\\
\InTextDef{20ex}{\override(\none,\overline{\mh\QM;})}{\Aux{mostSpecific}(\overline{\mh\QM;})}\\
\InTextDef{20ex}{\override(\method,\methods)}{\method}\\
\InTextWith{
\forall \method' \in \methods \ \
%\method'\in\{\mh\QM;,\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \},
\method\subtype\method'
}
\end{array}$

\noindent The definition $\Aux{mostSpecific}$ returns the most
specific method, that is a method whose type is the subtype of all the
others.  Since method subtyping is a partial ordering,
$\Aux{mostSpecific}$ may not be defined, this in turn forces us to rely on the last clause of $\override$; otherwise
the whole \mBody{} would be not defined for that specific $\m$.
Rule \rn{t-intf} relies on this behaviour.

>>>>

Defining \mBody{} is not trivial, and requires quite a lot of attention to the
specific model of Java interfaces, and to how it differs w.r.t. Java Class model.
$\mBody(\m,\C)$ denotes the actual method $\m$ (body included) that
interface $\C$ owns. The method can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.
\noindent$\begin{array}{l}
\InTextDef{15ex}{
\mBody(\m,\C_0)
}{
\override(\methods(\m),
\tops(\m,\Cs))
}\\
\InTextWith{\metaVar{IT}(\C_0) =
\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_n\ \oC\methods\ \cC \mbox{ and }
 \C\in\Cs \mbox{ if } \C_i\subtype\C, i\in 1..n

@@

\yanlin{I think the explaination for needed might be hard to understand (for
  reviewer). need revising.}
\paragraph{\tops{}} recovers from the interface table only the ``needed'' methods, that is,
the non-static ones that are not reachable by another, less specific superinterface.
Since the second parameter of \tops{} is a set, we can choose an arbitrary element to be $\C_0$.
In the definition we denote by $\Aux{originalMethod}(\m,\C)=\method$ the non-static method called $\m$ defined directly in $\C$.
%\hl{Another two functions $\super(\C_0)$ and $\get(\m,\C_0)$ are used for the definition. $\super(\C_0)$ collects the
%super interfaces of $\C_0$ recursively, and $\get(\m,\C_0)$ returns the non-static method declared in $\C_0$ with name $\m$, or $\emptyset$ otherwise.
%Finally, $\tops(\m,\C_0)$ returns the topmost methods $m$ from the super interfaces.}
Formally:

@@

%\paragraph{\shadow{}} chooses the most specific version of a method,
%that is the unique version available, or a conflicted version from a
%set of possibilities. In this case all the conflicting defintions are returned.
%We do not model overloading, so it is an error if multiple versions are available with different parameter types. Formally:
%
%\noindent
%$\begin{array}{l}
%\InTextDef{15ex}{
%\shadow()}{\none}\\
%\InTextDef{15ex}{\shadow(\method)}{\method}\\
%\InTextDef{15ex}{\shadow(\overline{\mh\QM;})}{\Aux{mostSpecific}(\overline{\mh\QM;})}\\
%\InTextDef{15ex}{\shadow(\methods)}{\methods}\\
%\InTextWith{\mbox{ the former case not applicable} }
%%\methods\mbox{ not of the form }\overline{\mh\QM;}
%%\mbox{ and }
%%\Aux{mostSpecific}(\methods)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}}\\
%\end{array}$
%
\paragraph{override} models how a method in an interface can
override an implementation in the superinterface, even in case of a
conflict. Note how the special value $\none$ is used, and how (the 4th
case) overriding can solve a conflict.
% The notation in the last case below
%means $\method'$ is not static, and is supertype of $\method$.

\noindent$\begin{array}{l}
\InTextDef{20ex}{\override(\none,\emptyset)}{\none}\\
\InTextDef{20ex}{\override(\method,\emptyset)}{\method}\\
\InTextDef{20ex}{\override(\none,\method)}{\method}\\
\InTextDef{20ex}{\override(\none,\overline{\mh\QM;})}{\Aux{mostSpecific}(\overline{\mh\QM;})}\\
\InTextDef{20ex}{\override(\method,\methods)}{\method}\\
\InTextWith{
\forall \method' \in \methods : 
%\method'\in\{\mh\QM;,\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \},
\method\subtype\method'
}
\end{array}$
\yanlin{added a colon}
\noindent The definition $\Aux{mostSpecific}$ returns the most
specific method whose type is the subtype of all the
others.  Since method subtyping is a partial ordering,
$\Aux{mostSpecific}$ may not be defined, this in turn forces us to rely on the last clause of $\override$; otherwise
the whole \mBody{} would not be defined for that specific $\m$.
Rule \rn{t-intf} relies on this behaviour.
